---
title: "Multiple Linear Regression Runs"
author: "Uyen Nguyen"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Loading in tidyverse and dataset
library(tidyverse)
library(GGally)
library(MASS)

# Raw diabetes set (768 rows by 9 variables)
diabetes <- as_tibble(read.csv('diabetes2.csv'))


# Change Outcome variable to factor with appropriate labels
diabetes$Outcome <- factor(diabetes$Outcome)
levels(diabetes$Outcome)
levels(diabetes$Outcome) <- c("No","Yes")
levels(diabetes$Outcome)

# Diabetes set without Insulin, Skin Thickness and complete cases only (724 rows by 7 variables)
diabetes2 <- diabetes

zero_unknown_cols <-c("Glucose", "BloodPressure", "SkinThickness",
                      "Insulin", "BMI", "DiabetesPedigreeFunction",
                      "Age")

diabetes2[,zero_unknown_cols] <- replace(diabetes2[,zero_unknown_cols], diabetes2[,zero_unknown_cols]==0,NA)

diabetes2 <- dplyr::select(diabetes2, -c('SkinThickness', 'Insulin'))
diabetes2 <- diabetes2[complete.cases(diabetes2),]
```


# Last iteration of MLR

## Scatterplots
```{r}
GGally::ggpairs(diabetes2)

# All of the potential explanatory variables have positive correlation to glucose, though none have a correlation coefficient above 27%.
```

## Full MLR with Pregnancies, Blood Pressure, BMI, DiabetesPedigeeFunction, and Age
```{r}
result <- lm(Glucose ~ Pregnancies + BloodPressure + BMI + DiabetesPedigreeFunction + Age, data = diabetes2)
summary(result)

# The full model is highly statistically significant in predicting glucose, judging by the F-test statstic.  
# However, one of its explanatory variables, Pregnencies, is not (t-test p-value of 71%). 
```

```{r}
yhat <- result$fitted.values
res <- result$residuals
diabetes2 <- data.frame(diabetes2, yhat, res)

ggplot(diabetes2, aes(x = yhat,y = res))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y",
       y="Residuals",
       title="Residual Plot from full MLR")
       
# The residual scatterplot shows non-constant variance, suggesting a transformation on the y-variable is needed. 
```


```{r}
boxcox(result)

# According to the Box-Cox plot, we should take the log of the y-variable (glucose). 
```


```{r}
acf(result$residuals)

# The ACF plot does not show any significant autocorrelation. 
```


```{r}
qqnorm(result$residuals)
qqline(result$residuals, col = "red")

# The QQ plot is not a perfect line, but it is close enough. 
```


```{r}
ystar <- log(diabetes2$Glucose)
diabetes2 <- data.frame(diabetes2, ystar)
```

## Final MLR with Blood Pressure, BMI, DiabetesPedigeeFunction, and Age
```{r}
resultLog <- lm(ystar ~ BloodPressure + BMI + DiabetesPedigreeFunction + Age, data = diabetes2)
summary(resultLog)

# Here we fit a log-transformed version of the rediced model (dropping Pregnencies). 
# The model is highly significant (p-value of F-test is nearly 0). 
# All of the individual explanatory variables are significant at the 95% confidence level. 
# Unfortunately, the adjusted R-squared of the model is only 12.24%, which means it only explains 12.24% of the variation in blood glucose. 
```


```{r}
yhatLog <- resultLog$fitted.values
resLog <- resultLog$residuals
diabetes2 <- data.frame(diabetes2, yhatLog, resLog)

ggplot(diabetes2, aes(x = yhatLog,y = resLog))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted log y",
       y="Residuals",
       title="Residual Plot from Reduced 3 log")
       
# The residual plot looks much better (mostly random). 
```


```{r}
boxcox(resultLog)

# The Box-Cox plot lines include 1, so no further transformation is necessary. 
```


```{r}
acf(resultLog$residuals)

# The ACF plot does not show any significant autocorrelation. 
```



```{r}
qqnorm(resultLog$residuals)
qqline(resultLog$residuals, col = "red")

# The QQ plot is not a perfect line, but it is close enough. 
```


## Our final model is statistically significant as a whole, and all of its explanatory variables are significant as well. 
## The final model satisifies all the assumptions of linear regression. 
## The only problem is the low adjusted R-squared, which reduces the practical usefulness of the model.
